Secure Composition of Robust and Optimising Compilers

### 2023-07-11 popl/main.tex
format

\begin{definition}[\glsfirst{sms}]\label{def:trace:smsdef}
  $
  \smssafe:=\left\{\trace \left|\begin{array}{rcl}
                                  \text{If }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\le_{\trace}\ev{Use\ \loc\ m;\sandboxtag;\securitytag'}, \text{ then }m<n
                                \end{array}\right.\right\}
  $
\end{definition}

@@

\begin{definition}[\glsfirst{ms}]\label{def:trace:msdef}
  $
  \mssafe:=\left\{\trace \left|\begin{array}{rcl}
                                 \trace\in\tmssafe & \text{ and } & \trace\in\smssafe
                               \end{array}\right.\right\}
  $
\end{definition}

@@

\begin{definition}[\glsfirst{scct}]\label{def:trace:scctdef}
  $
  \scctsafe:=\left\{\trace \left|\begin{array}{rcl}
                                   \trace=\hole{\cdot} & \text{ or } & \trace=\event[_{b}];\sandboxtag;\unlock\cdot\trace['] \wedge \trace[']\in\scctsafe
                                 \end{array}\right.\right\}
  $
\end{definition}

@@

\begin{definition}[\glsfirst{ms} and \gls{scct}]\label{def:trace:msscctdef}
  $
  \msscctsafe:=\mssafe\cap\scctsafe
  $
\end{definition}

@@

\begin{figure}[h!]
  \judgbox{\traceagree{\pi}{\trace}{\abstrace}}{,,Specification-level trace $\trace$ agrees with monitor-level trace $\abstrace$ with respect to property $\pi$.''}
  %
  \typerule{traceagree-empty}{}{
    \traceagree{\pi}{\hole{\cdot}}{\hole{\cdot}}
  }{traceagree-empty}
  %
  \typerule{traceagree-ign-L}{
    \traceagree{\pi}{\trace}{\abstrace}
  }{
    \traceagree{\pi}{\emptyevent\cdot\trace}{\abstrace}
  }{traceagree-ign-L}
  %
  \typerule{traceagree-ign-R}{
    \traceagree{\pi}{\trace}{\abstrace}
  }{
    \traceagree{\pi}{\trace}{\absev{\emptyevent}\cdot\abstrace}
  }{traceagree-ign-R}
  %
  \typerule{traceagree-cons}{
    \traceagree[]{\pi}{\event}{\absevent} &
    \traceagree{\pi}{\trace}{\abstrace}
  }{
    \traceagree{\pi}{\event\cdot\trace}{\absevent\cdot\abstrace}
  }{traceagree-cons}
  \caption{Trace-Agreement relation that equates specification-level traces with monitor-level traces.}\label{fig:stuttereq}
\end{figure}

Monitors enforce safety properties by accepting or rejecting traces, i.e., if it rejects a trace, the trace does not satisfy the property the monitor checks.
Since reasoning on monitors is easier than directly on just traces, this section presents a monitor for each of the previously shown safety properties (\Cref{subsec:propdefs}).
To lessen the burden when proving that a monitor accepts the trace of a program execution, each monitor uses a custom trace model that contains only the relevant information related to the property the monitor checks.
To go from specification traces $\trace$ to monitor-level traces $\abstrace$, each property $\pi$ has an associated event agreement relation $\traceagree[]{\pi}{\event}{\absevent}$.
\Cref{fig:stuttereq} shows how the event agreement is lifted to traces.
The trace agreement is the same for all properties $\pi$ up to the use of the event agreement in \Cref{tr:traceagree-cons}.
With agreements, this section defines monitor satisfaction for traces and then it proves that monitor satisfaction implies property satisfaction.
To this end, monitor satisfaction is defined as follows.
\bul{A specification trace $\trace$ monitor-satisfies property $\pi$} iff there exists a (final) monitor state $\monitor$ and an abstract trace $\abstrace$ such that the \iul{specification trace $\trace$ agrees with abstract trace $\abstrace$} and \oul{the initial monitor\footnote{In this paper, for all monitors, the initial monitor state is denoted as $\emptyset$.} can step to the (final) monitor state $\monitor$ with abstract trace $\abstrace$}.

@@

\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
\bul{Define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
Given a program \texttt{p}, its compiled counterpart is obtained by \rul{plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by \oul{plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
\begin{definition}[Upper Composition]
  \bul{$\cc{\src{L}+\obj{L}}{\trg{L}}$} $\isdef$ Given \texttt{p}, yield
  $\left\{\mbox{\begin{tabular}{c}
    {\rul{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
    \mbox{\oul{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
  \end{tabular}}\right.$
%
%
%  $\begin{cases}
%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
%                                                                          \end{cases}$
\end{definition}

@@

Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
\bul{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
Given a program $\src{p}$, its compiled counterpart is obtained by \rul{plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by \oul{plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, \pul{based on the internal decision}.
\begin{definition}[Lower Composition]
  \bul{$\cc{\src{L}}{\trg{L}+\obj{L}}$} $\isdef$ Given $L$ and $\src{p}$, yield
  $\left\{\mbox{\begin{tabular}{c}
    {\pul{if $L=\trg{L}$, then} \rul{$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
    \mbox{\pul{if $L=\obj{L}$, then} \oul{$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
  \end{tabular}}\right.$
%
%
%  $\begin{cases}
%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
%                                                                          \end{cases}$
\end{definition}

@@

This section compares existing work with the one presented in this paper.
To this end, work on robust preservation (\Cref{subsec:relw:seccomprtp}) and on other secure compiler criteria (\Cref{subsec:relw:seccompcrit}) is looked at.
The case study of this paper (\Cref{sec:casestud:defs,sec:casestud:rtp}) implements measures for ensuring \gls{ms} as well as \gls{cct} and these are compared with previous work as well (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs}).

@@

The robust preservation of properties as a compiler-level criterion has been analyzed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2022universal,patrignani2019survey}.
This work has made use of that previous work.
While there exists a composition theorem~\cite{abate2019jour} already, that theorem is not concerned with composing robustly safe compilers, but rather program components and contexts.
The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
% The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:compcomp}) as well as in an earlier version of this work~\cite{kruse2022csc}.
% However, they do not demonstrate the scalability of the approach by means of a case study.

@@

While this paper focuses on the robust preservation framework~\cite{abate2019jour}, other secure compilation criteria exist.
The survey on formal approaches to secure compilation~\cite{patrignani2019survey} discusses a broad spectrum already, while this section presents a very high-level overview.
The full abstraction approach~\cite{abadi1999fullabstraction} states that a compiler should preserve and reflect observational equivalence between source and target programs.
It was shown~\cite{abate2021faandrc} that fully abstract compilers robustly preserve program properties that are either trivial or meaningless.
As a mitigation for this, the authors presented a categorical approach based on maps of distributive laws~\cite{watanabe2002modl}, which they call many maps of distributive laws.
Maps of distributive laws have been investigated before as a possible secure compilation criterion~\cite{tsampas2020catsc}.
Other approaches are extensions of the compiler correctness criterion as discussed in other work~\cite{patterson2019next700} or the introduction of opaque observations~\cite{vu2021reconciling} to reconcile compiler optimisations with security.
Note that this work also presents secure compilers that are optimising, but contrary to the other~\cite{vu2021reconciling}, provides a formal account of these in the robust preservation framework.
% Lastly, the authors of this paper have presented ongoing work~\cite{patrignani2023blame} on a weaker robust preservation criteria based on the concept blame.

@@

Different mechanisms for memory safety exist that also consider the secure compilation domain, i.e., have an active attacker model.
For example, the ,,pointers as capabilities'' principle represents pointers as machine-level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of linear typing~\cite{morrisett2005L3}.
The approach of this paper also uses linear typing, but differs from $L^{3}$~\cite{morrisett2005L3} in the way that functions are not first-class.
Moreover, this paper considers an active attacker, while the work on $L^{3}$ only discusses whole programs and, thus, has no active attacker model.
The instrumentation to ensure memory safety that this paper presents is inspired by seminal work in the literature~\cite{nagarakatte2009soft}.
That work inserts bounds-checks in front of pointer-dereferences and, for this to work, inserts meta-data information on pointer creation.
It also works in a more advanced setting with structured fields accesses and also introduces a table-lookup for pointers that are stored in memory.
This paper only considers pointers to arrays of simple objects, i.e., there are no pointers to pointers or structures.
Of course, there are a lot of other approaches to memory-safety in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware-extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}, to name a few.
What differentiates this work from them is that this work uses known, compiler-based approaches to ensure memory-safety as a means to investigate secure compiler compositions.
This paper does not provide efficient memory-safety, but serves as a theoretical foundation for the secure compilation domain.

@@

The approach to cryptographic constant time in this paper is high-level, where a programming language exposes a way to switch the semantics to a data (operand) independent timing mode.
Since identifiers in $\Lscct$ are annotated with a secrecy tag, this approach is similar to others with information flow control.
For example, Vale~\cite{bond2017vale} uses Dafny to ensure constant-time assembly code, while Jasmin~\cite{almeida2017jasmin} makes use of the Coq proof assistant to reject non-constant-time programs.
CT-Wasm~\cite{watt2019ctwasm} enforces constant-timeness by means of a type system.
Different to the approach of this paper, these approaches necessitate that the programmer writes \gls{cct} code.
An approach to allow programmers to write more high-level code is CryptOpt~\cite{kuepper2023cryptopt}, which generates efficient target-code by means of a randomised search.
This paper abstracts over concrete mitigation strategies and simply assumes that there is a flag to switch to a cryptographic-constant time execution mode.
This can be realised by employing the FaCT~\cite{cauligi2019fact} compiler, which translates common non-constant time code patterns to be constant-time, and the data (object) independent timing execution mode of modern processors.

@@

This paper does a first step towards practical secure compilation chains by introducing a theoretical framework demonstrating that secure compilers compose.
The paper exercises this framework on a case-study that consists of an optimising compilation chain that is secure with respect to \gls{msscct}.
In future work, it would be interesting to investigate whether it is possible to provide {\em secure compiler combinators}, similarly to how parser combinators work.
%To this end, it may be necessary to extend existing frameworks for multi-language semantics.
It is also interesting to extend our case-study and verify all results of it in Coq.
This way, it is possible to obtain an executable, secure compiler.
However, the formalisation effort for backtranslation proofs is known to be enormous, so another research avenue may be to find better ways to (semi-)automate standard secure compilation proofs.

%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
% \begin{acks}
%   We would like to thank the anonymous reviewers for their feedback.
% \end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{main}

>>>>

\begin{definition}[\glsfirst{sms}]\label{def:trace:smsdef}
  $$
  \smssafe:=\left\{\trace \left|\begin{array}{rcl}
                                  \text{If }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\le_{\trace}\ev{Use\ \loc\ m;\sandboxtag;\securitytag'}, \text{ then }m<n
                                \end{array}\right.\right\}
  $$
\end{definition}

@@

\begin{definition}[\glsfirst{ms}]\label{def:trace:msdef}
  $$
  \mssafe:=\left\{\trace \left|\begin{array}{rcl}
                                 \trace\in\tmssafe & \text{ and } & \trace\in\smssafe
                               \end{array}\right.\right\}
  $$
\end{definition}

@@

\begin{definition}[\glsfirst{scct}]\label{def:trace:scctdef}
  $$
  \scctsafe:=\left\{\trace \left|\begin{array}{rcl}
                                   \trace=\hole{\cdot} & \text{ or } & \trace=\event[_{b}];\sandboxtag;\unlock\cdot\trace['] \wedge \trace[']\in\scctsafe
                                 \end{array}\right.\right\}
  $$
\end{definition}

@@

\begin{definition}[\glsfirst{ms} and \gls{scct}]\label{def:trace:msscctdef}
  $$
  \msscctsafe:=\mssafe\cap\scctsafe
  $$
\end{definition}
\MPin{
	why is this defined differently than MS? they should be defined in the same way
}

@@

Monitors enforce safety properties by accepting or rejecting traces, i.e., if it rejects a trace, the trace does not satisfy the property the monitor checks.
Since reasoning on monitors is easier than directly on just traces, this section presents a monitor for each of the previously shown safety properties (\Cref{subsec:propdefs}).
To lessen the burden when proving that a monitor accepts the trace of a program execution, each monitor uses a custom trace model that contains only the relevant information related to the property the monitor checks.
To go from specification traces $\trace$ to monitor-level traces $\abstrace$, each property $\pi$ has an associated event agreement relation $\traceagree[]{\pi}{\event}{\absevent}$.
\Cref{fig:stuttereq} shows how the event agreement is lifted to traces.
\begin{figure}[!ht]
  \judgbox{\traceagree{\pi}{\trace}{\abstrace}}{,,Specification-level trace $\trace$ agrees with monitor-level trace $\abstrace$ with respect to property $\pi$.''}
  %
  \typerule{traceagree-empty}{}{
    \traceagree{\pi}{\hole{\cdot}}{\hole{\cdot}}
  }{traceagree-empty}
  %
  \typerule{traceagree-ign-L}{
    \traceagree{\pi}{\trace}{\abstrace}
  }{
    \traceagree{\pi}{\emptyevent\cdot\trace}{\abstrace}
  }{traceagree-ign-L}
  %
  \typerule{traceagree-ign-R}{
    \traceagree{\pi}{\trace}{\abstrace}
  }{
    \traceagree{\pi}{\trace}{\absev{\emptyevent}\cdot\abstrace}
  }{traceagree-ign-R}
  %
  \typerule{traceagree-cons}{
    \traceagree[]{\pi}{\event}{\absevent} &
    \traceagree{\pi}{\trace}{\abstrace}
  }{
    \traceagree{\pi}{\event\cdot\trace}{\absevent\cdot\abstrace}
  }{traceagree-cons}
  \caption{Trace-Agreement relation that equates specification-level traces with monitor-level traces.}\label{fig:stuttereq}
\end{figure}
The trace agreement is the same for all properties $\pi$ up to the use of the event agreement in \Cref{tr:traceagree-cons}.
With agreements, this section defines monitor satisfaction for traces and then it proves that monitor satisfaction implies property satisfaction.
To this end, monitor satisfaction is defined as follows.
\bul{A specification trace $\trace$ monitor-satisfies property $\pi$} iff there exists a (final) monitor state $\monitor$ and an abstract trace $\abstrace$ such that the \iul{specification trace $\trace$ agrees with abstract trace $\abstrace$} and \oul{the initial monitor\footnote{In this paper, for all monitors, the initial monitor state is denoted as $\emptyset$.} can step to the (final) monitor state $\monitor$ with abstract trace $\abstrace$}.

@@

\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
\bul{Define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
Given a program \texttt{p}, its compiled counterpart is obtained by \rul{plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by \oul{plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
\begin{definition}[Upper Composition]
  \bul{$\cc{\src{L}+\obj{L}}{\trg{L}}$} $\isdef$ 
  % Given \texttt{p}, yield
  $\lambda \texttt{p}\ldotp$
  $\left\{\mbox{\begin{tabular}{c}
    {\rul{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
    \mbox{\oul{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
  \end{tabular}}\right.$
%
%
%  $\begin{cases}
%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
%                                                                          \end{cases}$
\end{definition}

@@

Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
\bul{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
Given a program $\src{p}$, its compiled counterpart is obtained by \rul{plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by \oul{plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, \pul{based on the internal decision}.
\begin{definition}[Lower Composition]
  \bul{$\cc{\src{L}}{\trg{L}+\obj{L}}$} $\isdef$ 
  % Given $L$ and $\src{p}$, yield
  $\lambda \src{p}, L\ldotp$
  $\left\{\mbox{\begin{tabular}{c}
    {\pul{if $L=\trg{L}$, then} \rul{$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
    \mbox{\pul{if $L=\obj{L}$, then} \oul{$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
  \end{tabular}}\right.$
%
%
%  $\begin{cases}
%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
%                                                                          \end{cases}$
\end{definition}

@@

This section discusses work on robust preservation\MP{
	robust compilation.
	NOT robust preservation
} (\Cref{subsec:relw:seccomprtp}) and on other secure compilation criteria (\Cref{subsec:relw:seccompcrit}).
Since the case study of \Cref{sec:casestud:defs,sec:casestud:rtp} implements measures for preserving \gls{ms} and \gls{cct}, this section then presents relevant related work as well (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs}).

@@

The robust preservation of properties as a compiler-level criterion has been analyzed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2019survey} and thus we build on that framework.
No existing work is concerned with composing robustly safe compilers.
\MP{perhaps say taht we believe this scales to different trace models too and differnet kinds of props, but the ocmposition of monitors becomes tricky}
The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.\MP{why}
% The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:compcomp}) as well as in an earlier version of this work~\cite{kruse2022csc}.
% However, they do not demonstrate the scalability of the approach by means of a case study.

@@

While this paper focuses on the robust preservation framework~\cite{abate2019jour}, other secure compilation criteria exist.
The survey on formal approaches to secure compilation~\cite{patrignani2019survey} discusses a broad spectrum already, while this section presents a very high-level overview.
Fully abstract compilation~\cite{abadi1999fullabstraction} states that a compiler should preserve and reflect observational equivalence between source and target programs.
It was shown~\cite{abate2021faandrc} that fully abstract compilers robustly preserve program properties that are either trivial or meaningless.
As a mitigation for this, the authors presented a categorical approach based on maps of distributive laws~\cite{watanabe2002modl}, which they call many maps of distributive laws.
Maps of distributive laws have been investigated before as a possible secure compilation criterion~\cite{tsampas2020catsc}.
Other approaches are extensions of the compiler correctness criterion as discussed in other work~\cite{patterson2019next700} or the introduction of opaque observations~\cite{vu2021reconciling} to reconcile compiler optimisations with security.
Note that this work also presents secure compilers that are optimising, but contrary to the other~\cite{vu2021reconciling}, provides a formal account of these in the robust preservation framework.
% Lastly, the authors of this paper have presented ongoing work~\cite{patrignani2023blame} on a weaker robust preservation criteria based on the concept blame.

@@

Different mechanisms for \MP{enforcing?}memory safety exist that also consider the secure compilation domain, i.e., have an active attacker model.
For example, the ,,pointers as capabilities'' principle represents pointers as machine-level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of linear typing~\cite{morrisett2005L3}.
The approach of this paper also uses linear typing, but differs from $L^{3}$~\cite{morrisett2005L3} in the way that functions are not first-class.
Moreover, this paper considers an active attacker, while the work on $L^{3}$ only discusses whole programs and, thus, has no active attacker model.
The instrumentation to ensure memory safety that this paper presents is inspired by seminal work in the literature~\cite{nagarakatte2009soft}.
That work inserts bounds-checks in front of pointer-dereferences and, for this to work, inserts meta-data information on pointer creation.
It\MP{what} also works in a more advanced setting with structured fields accesses and also introduces a table-lookup for pointers that are stored in memory.
This paper only considers pointers to arrays of simple objects\MP{NOT OBJECTS}, i.e., there are no pointers to pointers or structures.
Several other approaches to memory-safety exist in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware-extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}.
What differentiates this work from them is that this work uses known, compiler-based approaches to ensure memory-safety as a means to investigate secure compiler compositions.
This paper does not provide efficient memory-safety, but serves as a theoretical foundation for the secure compilation domain.

@@

The approach to \MP{preserving?enforcing?} cryptographic constant time in this paper is high-level, where a programming language exposes a way to switch the semantics to a data (operand) independent timing mode.
Since identifiers in $\Lscct$ are annotated with a secrecy tag, this approach is similar to others with information flow control.
For example, Vale~\cite{bond2017vale} uses Dafny to ensure constant-time assembly code, while Jasmin~\cite{almeida2017jasmin} makes use of the Coq proof assistant to reject non-constant-time programs.
CT-Wasm~\cite{watt2019ctwasm} enforces constant-timeness by means of a type system.
Different to the approach of this paper, these approaches necessitate that the programmer writes \gls{cct} code.
An approach to allow programmers to write more high-level code is CryptOpt~\cite{kuepper2023cryptopt}, which generates efficient target-code by means of a randomised search.
This paper abstracts over concrete mitigation strategies and simply assumes that there is a flag to switch to a cryptographic-constant time execution mode.
This can be realised by employing the FaCT~\cite{cauligi2019fact} compiler, which translates common non-constant time code patterns to be constant-time, and the data (object) independent timing execution mode of modern processors.

@@

This paper does a first step towards practical secure compilation chains by introducing a theoretical framework demonstrating that secure compilers compose.
The paper exercises this framework on a case-study that consists of an optimising compilation chain that is secure with respect to \gls{msscct}.
\MP{
	not an appropriate recap.
	rewrite summarising the paper
}

In future work, it would be interesting to investigate whether it is possible to provide {\em secure compiler combinators}, similarly to how parser combinators work.
%To this end, it may be necessary to extend existing frameworks for multi-language semantics.
It is also interesting to extend our case-study and verify all results of it in Coq.
This way, it is possible to obtain an executable, secure compiler.
However, the formalisation effort for backtranslation proofs is known to be enormous\MP{cit + details}, so another research avenue may be to find better ways to (semi-)automate standard secure compilation proofs.
\MPin{
	not particularly informative FW. i'd rather not have it as such
}

%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.

% \begin{acks}
%   We would like to thank the anonymous reviewers for their feedback.
% 	This work was partially supported by a gift from
		% the Italian Ministry of Education through funding for the Rita Levi Montalcini grant (call of 2019).
% \end{acks}

\newpage
%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{main}
